A Variant of G\"{o}del's Ontological Proof \\ in a Natural Deduction Calculus \\ (Draft)

### 2015-01-26 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Fix calculus
Update GodProof-ND.tex
updated proof

@@

\section{New Proof}

\subsection{Possibly, God Exists}

\begin{theorem}
\label{T1}
Positive properties are possibly exemplified:
$$
\all \varphi. [P(\varphi) \imp \pos \ex x.\varphi(x)]
$$
\end{theorem}
\begin{proof} \hfill
\begin{prooftree}
        \AXC{Axiom \ref{A2}} \dashedLine
        \UIC{$ \all \varphi. \all \psi.[(P(\varphi) \wedge \nec \all x.[\varphi(x) \imp \psi(x)]) \imp P(\psi)]$} \RightLabel{$\all_E$}
        \UIC{$ \all \psi.[(P(\rho) \wedge \nec \all x.[\rho(x) \imp \psi(x)]) \imp P(\psi)]$} \RightLabel{$\all_E$}
        \UIC{$(P(\rho) \wedge \nec \all x.[\rho(x) \imp \neg \rho(x)]) \imp P(\neg \rho)$} \doubleLine
        \UIC{$(P(\rho) \wedge \nec \all x.[\neg \rho(x)]) \imp P(\neg \rho)$}
                        \AXC{Axiom \ref{A1}} \dashedLine
                        \UIC{$\all \varphi.[ P(\neg \varphi) \biimp \neg P(\varphi) ]$} \RightLabel{$\all_E$}
                        \UIC{$ P(\neg \rho) \biimp \neg P(\rho) $} \doubleLine
                 \BIC{$ (P(\rho) \wedge \nec \all x.[\neg \rho(x)]) \imp \neg P(\rho) $} \doubleLine
                 \UIC{$ P(\rho) \imp \pos \ex x.\rho(x) $} \RightLabel{$\all_I$}
                 \UIC{$\all \varphi.[ P(\varphi) \imp \pos \ex x.\varphi(x) ] $}
\end{prooftree}
\end{proof}

\begin{definition}
\label{D1}
A \emph{God-like} being possesses all positive properties:
$$
G(x) \biimp \forall \varphi. [P(\varphi) \to \varphi(x)]
$$
\end{definition}

\begin{axiom}
\label{A3}
The property of being God-like is positive:
$$
P(G)
$$
\end{axiom}

\begin{corollary}
\label{C1}
Possibly, God exists:
$$
\pos \ex x. G(x)
$$
\end{corollary}
\begin{proof} \hfill
\begin{prooftree}
\AXC{Axiom \ref{A3}} \dashedLine
\UIC{$P(G)$}
                 \AXC{Theorem \ref{T1}} \dashedLine
                 \UIC{$\all \varphi.[ P(\varphi) \imp \pos \ex x.\varphi(x) ]$} \RightLabel{$\all_E $}
                 \UIC{$ P(G) \imp \pos \ex x.G(x) $} \RightLabel{$\imp_E$}
    \BIC{$\pos \ex x. G(x)$}
\end{prooftree}
\end{proof}

\subsection{Being God is an essence of any God}

\begin{axiom}
\label{A4}
Positive properties are necessarily positive:
$$
\all \varphi.[P(\varphi) \to \Box \; P(\varphi)]
$$
\end{axiom}

@@

\begin{theorem}
\label{T2}
Being God-like is an essence of any God-like being:
$$
\all y.[G(y) \imp \ess{G}{y}]
$$
\end{theorem}
\begin{proof}
Let the following derivation with the open assumption $G(x)$ be $\Pi_1[G(x)]$:

\begin{prooftree}
\AXC{$ \neg P(\psi)^1$}
       \AXC{Axiom \ref{A1}} \dashedLine
       \UIC{$\forall \varphi.(\neg P(\varphi)\imp P(\neg\varphi))$}\RightLabel{$\all_E$}
       \UIC{$\neg P(\psi)\imp P(\neg\psi)$}\RightLabel{$\imp_E$}
 \BIC{$P(\neg\psi)$}\RightLabel{}
                   \AXC{$G(x)$} \dottedLine\RightLabel{D\ref{D1}}
                 \UIC{$\forall \varphi .(P(\varphi)\imp \varphi(x))$}\RightLabel{$\forall_E$}
                      \UIC{$P(\neg \psi)\imp \neg \psi(x)$}\RightLabel{$\imp_E$}
           \BIC{$ \neg \psi(x)$}\RightLabel{$\imp_E$}
            \AXC{$ \psi(x)^2$} \RightLabel{$\imp_E$}
             \BIC{$\bot $}\RightLabel{$\imp_I^1$}
             \UIC{$\neg \neg P(\psi)$}\RightLabel{$\neg\neg_E$}
              \UIC{$ P(\psi)$}\RightLabel{$\imp_I^2$}
              \UIC{$ \psi(x)\imp P(\psi)$}
\end{prooftree}

\noindent
Let the following derivation with the open assumption $G(x)$ be $\Pi_2[G(x)]$:
\begin{prooftree}
       \AXC{$\psi(x)^3$}
                     \AXC{$\Pi_1[G(x)] $} \dashedLine
                \UIC{$\psi(x)\imp P(\psi)$}\RightLabel{$\imp_E$}
                       \BIC{$P(\psi)$}
                 \AXC{Axiom \ref{A4}} \dashedLine
                \UIC{$\all \varphi .(P(\varphi)\imp \Box P(\varphi))$}\RightLabel{$\all_E$}
                 \UIC{$P(\psi)\imp \Box P(\psi)$}\RightLabel{$\imp_E$}
           \BIC{$\Box P(\psi)$}\RightLabel{$\imp_I^3$}
           \UIC{$\psi(x)\imp\Box P(\psi)$}
\end{prooftree}

\noindent
Let the following derivation without open assumptions be $\Pi_3$:

\begin{prooftree}
       \AXC{$P(\psi)^4$}
                     \AXC{$G(x)^5$} \dottedLine\RightLabel{D\ref{D1}}
                \UIC{$\all\varphi .(P(\varphi)\imp \varphi(x))$}\RightLabel{$\all_E$}
                \UIC{$P(\psi)\imp \psi(x)$}\RightLabel{$\imp_E$}
                       \BIC{$\psi(x)$}\RightLabel{$\imp_I^5$}
                       \UIC{$G(x)\imp \psi(x)$}\RightLabel{$\all_I$}
                \UIC{$\all x .(G(x)\imp \psi(x))$} \RightLabel{$\imp_I^4$}
                \UIC{$P(\psi)\imp \forall x .(G(x)\imp \psi(x))$}
\end{prooftree}

\noindent
Let the following derivation with the open assumption $G(x)$ be $\Pi_4[G(x)]$:
\begin{prooftree}
       \AXC{$\psi(x)^6 $} \RightLabel{}
                     \AXC{$\Pi_2$} \dashedLine
                \UIC{$\psi(x)\imp\Box P(\psi)$}\RightLabel{$\imp_E$}
                \BIC{$\Box P(\psi)$}
                         \AXC{$\Box P(\psi)^7 $}\RightLabel{$\Box_E$}
                               \UIC{$P(\psi)$}
                                \AXC{$\Pi_3$} \dashedLine\RightLabel{}
                                       \UIC{$P(\psi)\imp \all x .(G(x)\imp \psi(x))$}\RightLabel{$\imp_E$}
                               \BIC{$\all x .(G(x)\imp \psi(x))$}\RightLabel{$\Box_I$}
                        \UIC{$\Box \all x .(G(x)\imp \psi(x))$} \RightLabel{$\imp_I^7$}
                        \UIC{$\Box P(\psi)\imp \Box \all x .(G(x)\imp \psi(x))$} \RightLabel{$\imp_E$}
                  \BIC{$\Box \all x .(G(x)\imp \psi(x))$}\RightLabel{$\imp_I^6$}
           \UIC{$\psi(x)\imp\Box \all x .(G(x)\imp \psi(x))$}
\end{prooftree}

\noindent
The use of the necessitation rule above is correct, because the only open assumption $\Box P(\psi)$ is boxed. In the derivation of Theorem \ref{T2} below, the assumption $G(x)$ in the subderivation $\Pi_4[G(x)^8]$ is discharged by the rule labeled $8$.

\begin{prooftree}
       \AXC{$G(x)^8 $} \RightLabel{}
                     \AXC{$\Pi_4[G(x)^8]$} \dashedLine
                \UIC{$\psi(x)\imp\Box \all x .(G(x)\imp \psi(x))$}\RightLabel{$\all_I$}
                \UIC{$\all \psi .(\psi(x)\imp\Box \all x .(G(x)\imp \psi(x)))$}\RightLabel{$\wedge_I$}
                \BIC{$G(x)\wedge \forall \psi .(\psi(x)\imp\Box \all x .(G(x)\imp \psi(x)))$}\dottedLine\RightLabel{D\ref{D2}}
                               \UIC{$\ess{G}{x}$} \RightLabel{$\imp_I^8$}
                        \UIC{$G(x)\imp \ess{G}{x}$} \RightLabel{$\all_I$}
                         \UIC{$\all y.[G(y) \imp \ess{G}{y}]$}
\end{prooftree}
\end{proof}

\subsection{If God's existence is possible, it is necessary}

\begin{definition}
\label{D3}
\emph{Necessary existence} of an individual is the necessary exemplification of all its essences:
$$
E(x) \biimp \all \varphi.[\ess{\varphi}{x} \imp \nec \ex y.\varphi(y)]
$$
\end{definition}

\begin{axiom}
\label{A5}
Necessary existence is a positive property:
$$
P(E)
$$
\end{axiom}

\begin{lemma}
\label{L1}
If there is a God, then necessarily there exists a God:
$$
\ex z. G(z) \imp \nec \ex x. G(x)
$$
\end{lemma}
\begin{proof} \hfill
\begin{prooftree}
\AXC{$ $} \RightLabel{1}
\UIC{$\ex z. G(z)$}\RightLabel{$\ex_E$}
\UIC{$G(g)$}
\end{prooftree}

\begin{prooftree}
\AXC{$ $} \dashedLine
\UIC{$G(g)$}
        \AXC{Theorem \ref{T2}} \dashedLine
        \UIC{$\all y.[G(y) \imp \ess{G}{y}]$}\RightLabel{$\all_E$}
        \UIC{$G(g) \imp \ess{G}{g}$}\RightLabel{$\imp_E$}
    \BIC{$\ess{G}{g}$}
                \AXC{Axiom \ref{A5}} \dashedLine
                \UIC{$P(E)$}
                        \AXC{$ $} \dashedLine
                        \UIC{$G(g)$} \dottedLine \RightLabel{D\ref{D1}}
                        \UIC{$\all \varphi.[P(\varphi) \imp \varphi(g)]$}\RightLabel{$\all_E$}
                        \UIC{$P(E) \imp E(g)$}\RightLabel{$\imp_E$}
                     \BIC{$E(g)$} \dottedLine\RightLabel{D\ref{D3}}
                     \UIC{$ \all \varphi.[\ess{\varphi}{g} \imp \nec \ex x.\varphi(x)] $}\RightLabel{$\all_E$}
                     \UIC{$ \ess{G}{g} \imp \nec \ex x. G(x) $}\RightLabel{$\imp_E$}
        \BIC{$\nec \ex x. G(x)$} \RightLabel{$\imp_I^1$}
        \UIC{$\ex z. G(z) \imp \nec \ex x. G(x)$}
\end{prooftree}
\end{proof}

\subsection{God exists}

ToDo: this is proven in a way that is slightly different from G\"odel's 1970.

\begin{theorem}
\label{T4}
God exists:
$$
\ex x. G(x)
$$
\end{theorem}

\noindent
\textbf{Formal proof:}

\begin{small}
\begin{prooftree}
\AXC{$ $} \RightLabel{1}
\UIC{$\nec\neg\nec \ex x. G(x)$}\RightLabel{$\nec_E$, axiom T}
\UIC{$\neg\nec \ex x. G(x)$}
\AXC{ Lemma \ref{L1} } \dashedLine
\UIC{$\neg\nec \ex x. G(x) \imp \neg\ex x. G(x)$}\RightLabel{$\imp_E$}
 \BIC{$ \neg\ex x. G(x)$} \RightLabel{$\nec_I$}
\UIC{$\nec \neg\ex x. G(x)$} \dottedLine
\UIC{$\neg \pos\ex x. G(x)$}
\end{prooftree}
\end{small}

\begin{small}
\begin{prooftree}
\AXC{$ $}\dashedLine
\UIC{$\neg \pos\ex x. G(x)$}
\AXC{ Corollary \ref{C1}} \dashedLine
\UIC{$ \pos\ex x. G(x)$} \RightLabel{$\imp_E$}
 \BIC{$ \bot$} \RightLabel{$\imp_I^1$}
 \UIC{$\neg\nec\neg\nec \ex x. G(x)$} \dottedLine
 \UIC{$\neg\nec\pos \neg \ex x. G(x)$}
\AXC{ Axiom B} \dashedLine
\UIC{$ \neg \ex x. G(x)\imp\nec\pos \neg \ex x. G(x)$} \doubleLine
\UIC{$ \neg\nec\pos \neg \ex x. G(x)\imp \neg \neg \ex x. G(x)$} \RightLabel{$\imp_E$}
 \BIC{$ \neg \neg \ex x. G(x)$}\RightLabel{$ \neg \neg$ E}
 \UIC{$ \ex x. G(x)$}
\end{prooftree}
\end{small}

Note that the last step is classical and we do not prove the existential statement by providing an object for which the statement holds.
This proof makes section \ref{GodsE} superflous and the use of axiom "M" unnecessary.

The system used contains the $\nec_E$-rule with the restriction that we have a $\nec_I$ below it. This is equivalent to modal sytem K that contains axiom K and necessitation rule N. We aslo use axiom B ($ A\imp\nec\pos A$). No other modal axioms are needed.

\subsection{Necessarily, God exists}

We can also prove that god exists necessarily in our system by simply introducing box on a theorem by rule N in modal sytem K. So we get:

\begin{corollary}
\label{T3}
Necessarily, God exists:
$$
\nec \ex x. G(x)
$$
\end{corollary}

\subsection{God exists}\label{GodsE}
This section is superfluous.

\begin{axiom}[M]
\label{M} What is necessary is the case:
$$
\all \varphi. [\nec \varphi \imp \varphi]
$$
\end{axiom}

\begin{corollary}
\label{C2}
There exists a God:
$$
\ex x. G(x)
$$
\end{corollary}
\begin{proof}\hfill
\begin{prooftree}
\AXC{Theorem \ref{T3}} \dashedLine
\UIC{$ \nec \ex x. G(x)$}
        \AXC{$ \all \varphi. [\nec \varphi \imp \varphi]$}
        \UIC{$\nec \ex x. G(x) \imp \ex x. G(x)$}
    \BIC{$\ex x. G(x)$}
\end{prooftree}
\end{proof}

@@

@@

>>>>

\section{A distribution principle and axiom B}

To prove G\"odel's ontological proof in natural deduction we will prove that a distribution principle is derivable and state Brouwer's reduction theorem.

\begin{lemma}\label{DP}
The distribution principle 
$$\nec (A\imp B)\imp(\pos A\imp \pos B)$$
is provable in our system K of modal logic. 
\end{lemma}

\begin{proof}\hfill

\begin{small}
\begin{prooftree}
\AXC{$ $} \dashedLine\RightLabel{2}
\UIC{$\nec (A\imp B)$}\RightLabel{$\nec_E$}
\UIC{$\omega: A\imp B$}
\AXC{$ $} \dashedLine\RightLabel{1}
\UIC{$\pos A$}\RightLabel{$\pos_E$}
\UIC{$\omega: A$}\RightLabel{$\imp_E$}
\BIC{$\omega: B$} \RightLabel{$\pos_I$}
\UIC{$\pos B$}\RightLabel{$\imp_I^1$}
\UIC{$\pos A\imp \pos B$}\RightLabel{$\imp_I^2$}
\UIC{$\nec (A\imp B)\imp(\pos A\imp \pos B)$}
\end{prooftree}
\end{small}

Brouwer's reduction theorem is a version of axiom B for modal logic and a property that enforces symmetry as a geometric property for the frame of a Kripke sematics for modal logic. \citep{negri}
%See the following for a discussion about axiom B: http://plato.stanford.edu/entries/logic-modal/

@@

\section{New Proof}
\setcounter{axiom}{0}
\setcounter{lemma}{0}
\setcounter{theorem}{0}
\setcounter{corollary}{0}
\subsection{Possibly, God Exists}
\begin{axiom}
\label{A1}
Either a property or its negation is positive, but not both:
Axiom schema, for all $\varphi$:
$$
P(\neg \varphi) \biimp \neg P(\varphi)
$$
\end{axiom}
\begin{axiom}
\label{A2}
A property necessarily implied by a positive property is positive:
Axiom schema, for all $\varphi$ and $\psi$: 
$$
(P(\varphi) \wedge \nec \all x.[\varphi(x) \imp \psi(x)]) \imp P(\psi)
$$
\end{axiom}
\begin{theorem}
\label{T1}
Positive properties are possibly exemplified:
Axiom schema, for all $\varphi$:
$$
P(\varphi) \imp \pos \ex x.\varphi(x)
$$
\end{theorem}
\begin{proof} \hfill

\begin{prooftree}

\AXC{$[P(\rho)]^3$} 
\AXC{$[\nec \neg \ex x.\rho(x) ]^2$} \RightLabel{$\nec_E$}
\UIC{$\neg \ex x.\rho(x) $} 
\AXC{$[\rho(x)]^1$} \RightLabel{$\ex_I$}
\UIC{$\ex x.\rho(x) $} \RightLabel{$\imp_E$}
\BIC{$ \bot $}\RightLabel{$\imp_I, 1$}
\UIC{$ \neg \rho(x) $} \RightLabel{$\imp_I$}
\UIC{$ \rho(x) \imp \neg \rho(x) $} \RightLabel{$\all_I$}
\UIC{$ \all x.(\rho(x) \imp \neg \rho(x)) $}  \RightLabel{$\nec_I$}
\UIC{$ \nec \all x.(\rho(x) \imp \neg \rho(x)) $}\RightLabel{$\wedge_I$}
\BIC{$ P(\rho) \wedge \nec \all x.[\rho(x) \imp \neg \rho(x)]$}
\AXC{Axiom \ref{A2} for $\rho$ and $\neg\rho$} \dashedLine
\UIC{$(P(\rho) \wedge \nec \all x.[\rho(x) \imp \neg \rho(x)]) \imp P(\neg \rho)$} \RightLabel{$\imp_E$}
\BIC{$P(\neg \rho)$} \RightLabel{$\imp_E$}
\end{prooftree}

\begin{prooftree}
\AXC{$ $} \dashedLine
\UIC{$P(\neg \rho)$} \RightLabel{$\imp_E$}
\AXC{Half of axiom \ref{A1}} \dashedLine
\UIC{$ P(\neg \rho) \imp \neg P(\rho) $}  \RightLabel{$\imp_E$}
\BIC{$\neg P(\rho) $} 
\AXC{$[P(\rho)]^3$} 
\BIC{$\bot$} \RightLabel{$\imp_I, 2$}
\UIC{$ \neg \nec \neg \ex x.\rho(x)$} \dashedLine \RightLabel{definition of $\pos$}
\UIC{$ \pos \ex x.\rho(x)$}  \RightLabel{$\imp_I, 3$}
\UIC{$ P(\rho) \imp \pos \ex x.\rho(x) $}

\end{prooftree}

\end{proof}
\begin{definition}
\label{D1}
A \emph{God-like} being possesses all positive properties:
$$
G(x) \biimp \forall \varphi. [P(\varphi) \to \varphi(x)]
$$
\end{definition}

\begin{axiom}
\label{A3}
The property of being God-like is positive:
$$
P(G)
$$
\end{axiom}
\begin{corollary}
\label{C1}
Possibly, God exists:
$$
\pos \ex x. G(x)
$$
\end{corollary}
\begin{proof} \hfill
\begin{prooftree}
\AXC{Axiom \ref{A3}} \dashedLine
\UIC{$P(G)$}
\AXC{Theorem \ref{T1} for G} \dashedLine
\UIC{$ P(G) \imp \pos \ex x.G(x) $} \RightLabel{$\imp_E$}
\BIC{$\pos \ex x. G(x)$}
\end{prooftree}
\end{proof}

\subsection{If God exists, then God necessarily exists}

@@

\begin{definition}
\label{D3}
\emph{Necessary existence} of an individual is the necessary exemplification of all its essences:
$$
E(x) \biimp \all \varphi.[\ess{\varphi}{x} \imp \nec \ex y.\varphi(y)]
$$
\end{definition}
\begin{axiom}
\label{A5}
Necessary existence is a positive property:
$$
P(E)
$$
\end{axiom}
\begin{lemma}
\label{L1}
If there is a God, then necessarily there exists a God:
$$
\ex z. G(z) \imp \nec \ex x. G(x)
$$
\end{lemma}

\begin{proof} \hfill
\begin{prooftree}
\AXC{$ $} \RightLabel{1}
\UIC{$\ex z. G(z)$}\RightLabel{$\ex_E$}
\UIC{$G(g)$}
\end{prooftree}

\begin{prooftree}

\AXC{$ $} \RightLabel{2}
\UIC{$\neg P(\psi)$} 
\AXC{Axiom \ref{A1}} \dashedLine
\UIC{$\neg P(\psi) \imp P(\neg \psi)$}\RightLabel{$\imp_E$}
\BIC{$P(\neg \psi)$}
\AXC{$G(g)$}  \dashedLine \RightLabel{D\ref{D1}}
\UIC{$\forall \varphi.(P(\varphi) \imp \varphi(g))$}\RightLabel{$\forall_E$}
\UIC{$P(\neg \psi) \imp \neg \psi(g)$}\RightLabel{$\imp_E$}
\BIC{$\neg \psi(g)$}
\AXC{$ $} \RightLabel{4}
\UIC{$\psi(g)$} \RightLabel{$\imp_E$}
\BIC{$\bot$}\RightLabel{$\imp_I^2$}
\UIC{$\neg\neg P(\psi)$}\RightLabel{$\neg\neg_E$}
\UIC{$P(\psi)$}
\end{prooftree}

\begin{prooftree}
\AXC{$P(\psi)$}
\AXC{$ $} \RightLabel{3}
\UIC{$G(y)$}  \dashedLine \RightLabel{D\ref{D1}}
\UIC{$\forall \varphi.(P(\varphi) \imp \varphi(g))$}\RightLabel{$\forall_E$}
\UIC{$P(\psi) \imp \psi(y)$}\RightLabel{$\imp_E$}
\BIC{$\psi(y)$} \RightLabel{$\imp_I^3$}
\UIC{$G(y) \imp \psi(y)$}\RightLabel{$\forall_I$}
\UIC{$\forall y.(G(y) \imp \psi(y))$}\RightLabel{$\nec_I$}
\UIC{$\nec \forall y.(G(y) \imp \psi(y))$}\RightLabel{$\imp_I^4$}
\UIC{$\psi(g) \imp \nec \forall y.(G(y) \imp \psi(y))$}\RightLabel{$\forall_I$}
\UIC{$\forall \psi.(\psi(g) \imp \nec \forall y.(G(y) \imp \psi(y)))$}\RightLabel{$\wedge_I$}
\AXC{$G(g)$}  
\BIC{$G(g) \wedge \forall \psi.(\psi(g) \imp \nec \forall y.(G(y) \imp \psi(y)))$}\dashedLine \RightLabel{D\ref{D2}}
\UIC{$\ess{G}{g}$}
\AXC{Axiom \ref{A5}} \dashedLine
\UIC{$P(E)$}\dottedLine \RightLabel{D\ref{D3}}
\AXC{$G(g)$}\dashedLine \RightLabel{D\ref{D1}}
\UIC{$\forall \varphi.(P(\varphi) \imp \varphi(g))$}\RightLabel{$\forall_E$}
\UIC{$P(E) \imp E(g)$}\RightLabel{$\imp_E$}
\BIC{$E(g)$}\dottedLine \RightLabel{D\ref{D3}}
\UIC{$ \all \varphi.[\ess{\varphi}{g} \imp \nec \ex x.\varphi(x)] $}\RightLabel{$\all_E$}
\UIC{$ \ess{G}{g} \imp \nec \ex x. G(x) $}\RightLabel{$\imp_E$}
\BIC{$\nec \ex x. G(x)$} \RightLabel{$\imp_I^1$}
\UIC{$\ex z. G(z) \imp \nec \ex x. G(x)$}
\end{prooftree}
\end{proof}

\subsection{God necessarily exists}

ToDo: this is proven in a way that is slightly different from G\"odel's 1970.
\begin{theorem}
\label{T4}
God necessarily exists:
$$
\nec \ex x. G(x)
$$
\end{theorem}
\begin{proof}

\begin{small}
\begin{prooftree}
\AXC{ Distribution principle \ref{DP} } \dashedLine
\UIC{$\nec [\ex x. G(x)\imp \nec \ex x. G(x)]\imp [\pos \ex x. G(x)\imp \pos\nec \ex x. G(x)]$}
\AXC{ Lemma \ref{L1} } \dashedLine
\UIC{$\ex x. G(x)\imp \nec \ex x. G(x)$}\RightLabel{$\nec_I$}
\UIC{$\nec[\ex x. G(x)\imp \nec \ex x. G(x)]$}\RightLabel{$\imp_E$}
\BIC{$\pos \ex x. G(x)\imp \pos\nec \ex x. G(x)$}
\end{prooftree}
\end{small}
\begin{small}
\begin{prooftree}
\AXC{$[\pos \ex x. G(x)\imp \pos\nec \ex x. G(x)]$}
\AXC{ Theorem \ref{T1} } \dashedLine
\UIC{$\pos \ex x. G(x)$}\RightLabel{$\imp_E$}
\BIC{$\pos\nec \ex x. G(x)$}
\AXC{ Brouwer's reduction theorem \ref{***} } \dashedLine
\UIC{$\pos\nec \ex x. G(x) \imp \ex x. G(x)$}\RightLabel{$\imp_E$}
\BIC{$ \ex x. G(x)$} \RightLabel{$\nec_I$}
\UIC{$\nec \ex x. G(x)$} 
\end{prooftree}
\end{small}

@@

\bibitem[{\itshape Anderson (1990)}]{and90}
Anderson, C. A. 1990. {\itshape Some Emendations of G\"odel's Ontological Proof}. Faith and Philosophy, Vol. 7, Issue 3, pp. 291-303.

@@

\bibitem[{\itshape Negri (2005)}]{negri}
Negri, S. 2005. {\itshape Proof analysis in modal logic}. J. Philosophical Logic, vol. 34, pp. 507--544.
